---
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Joint Density Functions

We can visualize part of the following joint density function using as a 3D surface.

$$
f(x, y) = 6e^{-2x- 3y} \quad \quad (x, y > 0)
$$

```{r}
f <- function(x, y) {
  6 * exp(-2 * x - 3 * y)
}

dx <- .1
dy <- .1
x <- seq(0, 3, by = dx)
y <- seq(0, 3, by = dy)
z <- outer(x, y, f)
persp(x, y, z, theta = 25, phi = 25)
```

Note that the function is defined or all $x, y > 0$, but most of the density is in this region around the origin.


## Computing Joint Probabilities

We represent the following probability graphically:

$$
P(X > 1, Y > .5) = \int_{y = .5}^{\infty} \int_{x = 1}^{\infty} f(x, y) dx, dy
$$


```{r}
z2 <- z
z2[1:10,] <- 0
z2[, 1:5] <- 0

persp(x, y, z2, theta = 25, phi = 25, zlim = c(0, max(z)))
```

### Computing the probability analytically

On board.

### Computing the probability computationally

We approximate the integral by discretizing the continuous volume into a series
of rectangles of height $f(x, y)$ and base area $dxdy$ and summing the probabilities.

$$
P(X > 1, Y > .5) \approx \sum_{x,y\textrm{ in grid}} f(x, y) dx, dy
$$


```{r}
sum(z2 * dx * dy)
```


### Practice

We can visualize $P(X > 1 \textrm{ or } Y > .5)$,

```{r}
z3 <- z
z3[1:10, 1:5] <- 0
persp(x, y, z3, theta = 25, phi = 25, zlim = c(0, max(z)))
```

then approximate it computationally.

```{r}
sum(z3 * dx * dy)
```

Or we can visualize the complement, $P(X < 1 \textrm{ or } Y < .5)$,

```{r}
z4 <- z
z4[11:31, ] <- 0
z4[, 6:31] <- 0
persp(x, y, z4, theta = 25, phi = 25, zlim = c(0, max(z)))
```

and compute 1 minus its volume.

```{r}
1 - sum(z4 * dx * dy)
```

The fact that these methods aren't in close agreement is likely due to the disagreement
between the discrete and continuous densities around the origin.

## Marginal Probabilities

We can visualize the process of integrating out $x$ to get $f_Y(y)$.

```{r}
fy <- colSums(z)
plot(y, fy, pch = 16)
lines(y, fy, col = "tomato", lwd = 2)
```

And equivalently for $f_X(x)$
```{r}
fx <- rowSums(z)
plot(y, fy, pch = 16)
lines(y, fy, col = "tomato", lwd = 2)
points(x, fx, pch = 16)
lines(x, fx, col = "steelblue", lwd = 2)
```


